\newcommand{\hmwkTitle}{Homework 4}
\newcommand{\hmwkDueDate}{Tuesday, May 19, 2015}
\newcommand{\hmwkHeaderTitle}{ICCS206 Discrete Mathematic: Homework 4}
\input{hw_template.tex}

\section*{Problem 1}
Let $F_n$ be the $n$-th Fibonacci number defined by 

\[
F_1=1, F_2=1, F_n=F_{n-1}+F_{n-2}
\] 

Show that

\[
F_n = \frac{1}{\sqrt{5}}\left(\left(\frac{1+\sqrt{5}}{2}\right)^n - \left(\frac{1-\sqrt{5}}{2}\right)^n\right)
\]

\medskip
\solution

\begin{proof}
  as the nth index of fibonacci sequence is defined as $F_n = F_{n-1}+F_{n-
    2}$, it's a linear recurrence, and we will try viewed T(n) as $x^n$ (some
terms to the power of n) --- this will help us find the closed form.

  \[T(n) = x^n\]

  thus

  \begin{align*}
  T(n) &= T(n-1) + T(n-2) \\
  x^n &= x^{n-1} + x^{n-2}
  \end{align*}

  now, we divide those newly created terms by $x^{n-2}$, gives

  \[x^2 - x - 1 = 0\]

  using the quadratic formula on the characteristic equation, we can derived
  x as

  \[
  \begin{aligned}
  x &= \frac{1 \pm \sqrt{1^2 - 4}}{2} \\
  & = \frac{1 + \sqrt{5}}{2}, \frac{1 - \sqrt{5}}{2}
  \end{aligned}
  \]

  it's graspable by example that we cannot just pick one of the two solution
  (in this recurrence case), as the constant doesn't adds up: $F_1 = 1 =
  cx^1, F_2 = cx^2$, this along with a few more cases hinted (I still don't
  understand linear homogeneous recurrence fully) that we should use both
  solution of x, by the linear combination property:

  \[(A(n), B(n)) \rightarrow (c_1 A(n) + c_2 B(n)) && \quad \text{for solution to linear system}\]

  and, since we were given $F_1 = 1$ and $F_2 = 1$

  \[
  \begin{aligned}
    T(1) = 1 &= c_1 \left(\frac{1 + \sqrt{5}}{2}\right) + c_2 \left(\frac{1 - \sqrt{5}}
  {2}\right)
  \end{aligned}
  \]

  and

  \[
  \begin{aligned}
    T(2) = 1 &= c_1 \left(\frac{1 + \sqrt{5}}{2}\right)^2 + c_2 \left(\frac{1 -
  \sqrt{5}}{2}\right)^2
  \end{aligned}
  \]

  to solve the two linear system, we let

    \[\left(\frac{1 + \sqrt{5}}{2}\right)^2 = \left(\frac{1 + \sqrt{5}}{2}\right) + 1 && \quad \text{as } x^2 - x - 1 = 0 \text{ from characteristic equation}
    \]

  and 

    \[\left(\frac{1 - \sqrt{5}}{2}\right)^2 = \left(\frac{1 - \sqrt{5}}{2}\right) + 1
    \]

    we'll get

\[
  \begin{aligned}
    1 &= c_1 \left(\frac{1 + \sqrt{5}}{2}+ 1\right) + c_2 \left(\frac{1 - \sqrt{5}}{2} + 1\right) 
  \end{aligned}
  \]

  which could be rearranged as

\[
  \begin{aligned}
    1 &= \left(c_1 \frac{1 + \sqrt{5}}{2} + c_2 \frac{1 - \sqrt{5}}{2}\right) + \left(c_1 + c_2\right) 
  \end{aligned}
  \]

  which the first term is equivalent to $T(1)$, so 

\[
  \begin{aligned}
    1 &= 1 + \left(c_1 + c_2\right) 
  \end{aligned}
  \]

  \[c_1 = -c_2\]


  then, substitude $c_2$ into $T(1)$

  \[
  \begin{aligned}
    T(1) = 1 &= c_1 \left(\frac{1 + \sqrt{5}}{2}\right) - c_1 \left(\frac{1 - \sqrt{5}}
  {2}\right)
  \end{aligned}
  \]

  gives

  \[c_1 = \frac{1}{\left(\frac{1 + \sqrt{5}}{2}\right) - \left(\frac{1 - \sqrt{5}}{2}\right)}\]

  finally, we got $c_1 = \frac{1}{\sqrt(5)}\text{, and } c_2 = -c_1 = -\frac{1}{\sqrt(5)}$; therefore \boxed{$F_n = \frac{1}{\sqrt{5}} \left(\left(\frac{1 + \sqrt{5}}{2}\right)^n - \left(\frac{1 - \sqrt{5}}{2}\right)^n\right)$}
  \end{proof}

\vspace{1em}
\hrulefill

\section*{Problem 2}
You bring 5 cola yoyos and 6 grape yoyos for 10 students. Students pick two yoyos at random:
\begin{enumerate}[label=\alph*.]
    \item Two grapes: eat one grape, return the other.
    \item Two colas: eat one cola, return the other.
    \item One of each: eat the grape, return the cola.
\end{enumerate}
The last yoyo in the bag goes to you. Which flavor can you guarantee for yourself, and why?

\medskip
\solution

\begin{proof}
notate the amount of grape yoyos with $G$ and coke yoyos with
$C$, the question provide 3 cases as
\begin{enumerate}[label={}]
\item case 1: grape - grape $\mapsto G - 1$
\item case 2: grape - coke $\mapsto G - 1$
\item case 3: cola - cola $\mapsto C - 1$
\end{enumerate}
we ignore the first and second case, only focus on amount of
cola yoyos remaining, we state our invarient as
\[
I \equiv \text{[ }C \equiv 1 \pmod 2 \text{ ]}
\]
initially, we got $C = 5$
\[
5 \equiv 1 \pmod 2
\]
think of it as for the case of cola - cola, you took 2
cola yoyos out, eat one, put one back, so $C - 2 + 1 \ge 0
\rightarrow \inf(C) = 1$ --- that's when our loop invariant
terminates --- at $C = 1$, $G = 0$, therefore that one
remaining yoyo is cola flavour.
\end{proof}

\vspace{1em}
\hrulefill

\section*{Problem 3}
Start with the numbers $1,2,3,4,5,6$. Repeatedly pick two numbers $(a,b)$, remove them, and insert $|a-b|$. Show that the final remaining number is always odd.

\medskip
\solution

\begin{proof}
we state our invariant as
\[
I \equiv \left[\sum_{i=0}^n a_i \equiv 1 \pmod 2\right]
\]
where $a_i$ is a member of list $a$, with length $n$.
initially
\[
\sum_{i=1}^6 i = \frac{6 \times 7}{2} = 21;\quad 21 \bmod 2
= 1
\]
the invariant holds, now for each iteration, notate our first
summation as $S_0 = 21 \in \mathbb{O}$, following $S$ will be
\[
S \bmod 2 = (S_0 - a_i - a_j + |a_i - a_j|) \bmod 2 \qquad
\text{when } 1 \le i, j \le 6
\]
it's true that
\[
(a_i + a_j) \equiv (a_i - a_j) \equiv |a_i - a_j| \pmod 2
\]

thus

\[
\begin{aligned}
S \bmod 2 &= (S_0 - (a_i + a_j) + |a_i - a_j|) \bmod 2\\
&= (S_0 - (a_i + a_j) + (a_i + a_j)) \bmod 2\\
&\equiv S_0 \pmod 2
\end{aligned}
\]
the result remain an odd number in every iteration, our
invariant holds during the maintenance phase, and it'll hold
after the loop termination, thus the final remaining number
is always odd.
\end{proof}

\vspace{1em}
\hrulefill

\section*{Problem 4}
A circle is divided into six sectors labeled $1,0,1,0,0,0$. You may pick two neighboring sectors and increase both numbers by 1. Can you make all six numbers equal?

\medskip
\solution

\begin{proof}
I initially tried the modulo and the summation approach, it
didn't work, so \textbf{this proof is hinted by codex-cli}.
we'd want those adjacent terms, that got increment by 1 to
cancle out, while preserving the invariant, so

\[
    I \equiv [x_1 - x_2 + x_3 - x_4 + x_5 - x_6 \equiv 2 \pmod 3]
\]

initially, we got $(1 - 0 + 1 + 0) \bmod 3 = 2$, the invariant
holds. for our maintenance phase, by adding 1 to adjacent
entry, there's a total of 6 combination: $(x_1, x_2)$, $(x_2,
x_3)$, $(x_3, x_4)$, $(x_4, x_5)$, $(x_5, x_6)$, $(x_6,
x_1)$ --- as the sign of $x_j$ and $x_{(j + 1) \bmod 6}$
is opposite, and there should be no negative members, as we
started by the lowest base number of 0, then increments.

\[
\begin{aligned}
x'_j + (-x'_{j+1}) &\equiv (x_j + 1) + (-(x_{j+1} + 1)) \pmod
3\\
&= x_j - x_{j+1} \pmod 3
\end{aligned}
\]

thus, the invariant holds in every iteration, now for the
claim of equal numbers on all entry, as there're 6 entries

\[
6n \not\equiv x_1 - x_2 + x_3 - x_4 + x_5 - x_6 \equiv 2
\pmod 3
\]

and therefore, it's impossible to achieve all equal entries,
and the claim is false.
\end{proof}

\vspace{1em}
\hrulefill

\section*{Problem 5}
On an $8 \times 8$ chessboard with standard coloring you may flip the colors of any entire row or column (all squares in that row/column). Can you reach a configuration with exactly one black square?

\section*{Problem 6}
An island has 13 blue, 15 red, and 17 green pokemons. When two different-colored pokemons meet, they both transform into the third color (e.g., blue + red $\rightarrow$ two greens). Show that it is impossible to reach a state with only one color remaining. (Hint: consider counts modulo 3.)

\end{document}
