\newcommand{\hmwkTitle}{Homework 6}
\newcommand{\hmwkDueDate}{Sunday, March 6, 2016}
\newcommand{\hmwkHeaderTitle}{ICCS206 Discrete Mathematic: Homework 6}
\input{hw_template.tex}

\section*{Problem 1: Poker Hands}
For a standard 52-card deck (4 suits, 13 ranks), count the number of 5-card hands of each type. Provide reasoning for each count.
\begin{enumerate}[label=\arabic*.]
    \item Total 5-card hands.
    \item Royal straight flush.
    \item Four-of-a-kind.
    \item Full house.
    \item Straight flush (non-royal).
    \item Flush (not straight flush or royal).
    \item Straight (not straight flush or royal).
    \item Three-of-a-kind (not full house).
    \item Two-pair.
    \item One-pair (not full house).
    \item High card (none of the above). Provide both an easy and a ``cool'' method.
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item $\dbinom{52}{5}$
    \item 4
    \item $13 \cdot 48$
    \item $(13 \cdot \dbinom{4}{3}) \cdot (12 \cdot \dbinom{4}{2})
    \item $9 \cdot 4$
    \item $4 \cdot (\dbinom{13}{5} - 10)$
    \item $9 \cdot (4^5 - 4) - 1$
    \item $13 \cdot \dbinom{4}{3} \cdot 4^2 \cdot \dbinom{12}{2}$
    \item $13 \cdot \dbinom{4}{2} \cdot 12 \cdot \dbinom{4}{2} \cdot (52 - 4 - 4)$
    \item $13 \cdot \dbinom{4}{2} \cdot \dbinom{12}{3} \cdot {4^3}$
    \item $\frac{\dbinom{13}{1} \cdot 4 \cdot\dbinom{12}{1} \cdot 4 \cdot \dbinom{11}{1} \cdot 4 \cdot \dbinom{10}{1} \cdot 4 \cdot \dbinom{9}{1} \cdot 4}{5!} - (9 \cdot 4^5) - (4 \cdot \dbinom{13}{5}) + (10 \cdot 4)$ \\
        $\simeq \frac{\text{(all cases)}}{\text{(we don't count permutation, count the thing as a set)}} - \text{(straight)} - \text{(flushes)} + \text{(straight flush)}$
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 2}
How many integers from 1 to 1000 are divisible by 2 or 5 or 7?

\medskip
\solution

using the 3 ways inclusion-exclusion diagram, we got 

\[|A \cup B \cup C| = |A| + |B| + |C| - |A \cap B| - |A \cap C| - |B \cap C| + |A \cap B \cap C| \]

which gives

\[ \left\lfloor \frac{1000}{2} \right\rfloor + \left\lfloor \frac{1000}{5} \right\rfloor + \left\lfloor \frac{1000}{7} \right\rfloor - \left\lfloor \frac{1000}{10} \right\rfloor - \left\lfloor \frac{1000}{14} \right\rfloor - \left\lfloor \frac{1000}{35} \right\rfloor + \left\lfloor \frac{1000}{70} \right\rfloor\]

\vspace{1em}
\hrulefill

\section*{Problem 3}
Decide whether each function $A \to B$ is injective, surjective, or bijective.
\begin{enumerate}[label=\arabic*.]
    \item $A = \{a,b\}$, $B = \{1,2\}$, $f(a)=1$, $f(b)=1$.
    \item $A = \{a,b\}$, $B = \{1,2,3\}$, $g(a)=1$, $g(b)=2$.
    \item $A = \{a,b,c\}$, $B = \{1,2\}$, $h(a)=1$, $h(b)=2$, $h(c)=1$.
    \item $A = \{a,b\}$, $B = \{1,2\}$, $q(a)=1$, $q(b)=2$.
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item neither injective, nor surjective
    \item injective only
    \item surjective only
    \item injective and surjective, so it's bijective as well
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 4}
Find a combinatorial proof of
\[
\sum_{i=0}^n \binom{n}{i}\binom{n}{n-i} = \binom{2n}{n}.
\]

\medskip
\solution

\begin{proof}
say we have n minors, and n wealthy guys on an little Saint James island. we can pick n humans from the total of 2n human by doing $\dbinom{2n}{n}$; or we could firstly pick i minors --- that's $\dbinom{n}{i}$, then pick the remaining $n - i$ humans, which identifies as a wealthy pedophiles --- that's $\dbinom{n}{n-i}$. as these two combination result are in-sequence to each other, we use the product rule, and to accounted for all combinations, $i = 0, 1, 2, ..., n$ --- which gives $\sum_{i=0}^{n} \dbinom{n}{i} \dbinom{n}{n-i}$

thus 

\[
\sum_{i=0}^n \binom{n}{i}\binom{n}{n-i} = \binom{2n}{n}.
\]

\remark we'll have to assume that all the child victims and those who had their name on the epstein file are all alike, for this combination analogy to work. 

\remark I don't think I should make the proof morally-related, I'm too obcessed with this, as my childhood idol, Stephen Hawking, is on the list as well; really hope it just research fundings. 

\end{proof}

\vspace{1em}
\hrulefill

\section*{Problem 5: Jubmoo}
Four players each receive 13 cards from a shuffled 52-card deck.
\begin{enumerate}[label=\arabic*.]
    \item How many total deals are possible?
    \item How many deals give Tow$+$ no hearts? (Others may also have none.)
    \item How many deals give Tow$+$ and Boss no hearts?
    \item How many deals give Tow$+$, Boss, and Majeed no hearts?
    \item How many deals give all four players no hearts?
    \item How many deals result in at least one player with no hearts? (Use inclusion-exclusion.)
    \item What is the probability of needing to redeal (at least one player has no hearts)? Provide a numerical value (approximately 5\%). 
\end{enumerate}


\medskip
\solution

\begin{enumerate}
    \item $\frac{52!}{13! \cdot 13! \cdot 13! \cdot 13!}$
    \item $\dbinom{39}{13} \cdot \frac{39!}{13! \cdot 13! \cdot 13!} \simeq \text{(no hearts for Tow+)} \cdot \text{(pick other three as normal)}$ 
    \item $\dbinom{39}{13} \cdot \dbinom{26}{13} \cdot \frac{26!}{13! \cdot 13!} \simeq \text{(no hearts for Tow+)} \cdot \text{no hearts for Boss} \cdot \text{(pick other two as normal)}$ 
    \item $\dbinom{39}{13} \cdot \dbinom{26}{13} \cdot \dbinom{13}{13} \cdot \frac{13!}{13!} \simeq \text{(no hearts for Tow+)} \cdot \text{no hearts for Boss} \cdot \text{no hearts for Majeed} \cdot \text{(pick other guy as normal)}$ 
    \item impossible, as for four players to not get any hearts, there must be $13 \cdot 4 = 52$ non-heart card in the deck, but a deck of cards already contain 52 cards, so that's not possible, as there must also be 13 heart cards in the deck. 
    \item let $A_1, A_2, A_3 \text{ and } A_4$ be each of the four players, as they're the same

        \[
            |A_1 \cup A_2 \cup A_3 \cup A_4| = \dbinom{4}{1}|A_i| + \dbinom{4}{2}|A_i \cap A_j| - \dbinom{4}{3}|A_i \cap A_j \cap A_k| + \dbinom{4}{4}{|A_i \cap A_j \cap A_k \cap A_l|}
        \]

        as $0 \le i, j, k, l \le 4$, and we get each of the terms from the previous: (5.2), (5.3), (5.4) and (5.5) for first, second, thrid and fourth term accordingly. 

    \item the probability they had to redeal the card is $\frac{P_{\text{someone has no heart card}}}{P_{\text{all}}}$ which is

        \begin{align*}
            \frac{(5.6)}{(5.1)} &= \frac{\binom{4}{1}\binom{39}{13}\frac{39!}{13!\cdot13!
        \cdot13!} - \binom{4}{2}\binom{39}{13}\binom{26}{13}\frac{26!}{13!\cdot13!}
        + \binom{4}{3}\binom{39}{13}\binom{26}{13}\binom{13}{13}\frac{13!}{13!}}
        {\frac{52!}{13!\cdot13!\cdot13!\cdot13!}} \\
        &= \frac{1621364909}{31750677980} \approx \boxed{0.05106552086923342}
        \end{align*}

\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 6: Mixed-Up Deck}
Three identical decks are mixed (156 cards total). Cards of the same rank and suit are indistinguishable. How many distinct arrangements are possible?

\medskip
\solution

as there're 156 cards, that's $156!$ arrangements, but since there'll exist 3 copy of 52 same exact cards (as this group of cards are built up from 3 full deck of cards), we'll have to remove all that repeatitive: 3 exact copies, 52 unique cards --- so that's \boxed{$156! \cdot \frac{1}{(3!)^{52}}$}

\vspace{1em}
\hrulefill

\section*{Problem 7: Grid-Walking Robot}
A robot moves only up or right along grid lines from $(0,0)$ to $(20,10)$.
\begin{enumerate}[label=\arabic*.]
    \item How many paths reach $(20,10)$?
    \item With a checkpoint at $(10,5)$, how many paths go through the checkpoint before reaching $(20,10)$?
    \item If there's also a pitfall at $(4,3)$ that must be avoided, how many paths go through $(10,5)$ and avoid $(4,3)$?
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item as the treasure is placed at (20, 10), the robot will have to do 20 right move, and 10 up move. if we list all the moves the robot makes until it reaches the treasure --- that'll be a list of length 30; then pick the slot that the robot will move up --- so 10 slots --- and that's  \boxed{$\dbinom{30}{10}$} paths.
    \item do it just like the first one, but as you have to pass the checkpoint, and you couldn't go backwards, it's traveling from (0, 0) to (10, 5), then (10, 5) to (20, 10), and as these two are sequencial, we use product rule --- which is \boxed{$\dbinom{15}{5} \cdot \dbinom{15}{5}$}
    \item just like the second one, subtract all path that pass the pitfall: (0, 0) to (4, 3) and (4, 3) to (10, 5); and we have to consider the sequencial paths from (10, 5) to (20, 10) as well --- which gives \boxed{$\dbinom{15}{5} \cdot \dbinom{15}{5} - \dbinom{7}{3}\dbinom{8}{2}\dbinom{15}{5}$}
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 8: Tower of Hanoi}
With three poles and 10 distinct disks (larger disks cannot sit on smaller disks):
\begin{enumerate}[label=\arabic*.]
    \item How many valid configurations exist when poles are distinct?
    \item How many when poles are identical? (Answer must be an integer.)
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item sort the disc, then let the largest disc choose first, which there's three poles to choose for each of the disc --- so \boxed{$3^{10}$}
    \item $\frac{3^{10}}{3!}$
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 9: Books on Shelves}
How many ways can $n$ books be placed on $k$ distinguishable shelves?
\begin{enumerate}[label=\arabic*.]
    \item Books are indistinguishable copies of one title.
    \item All books are distinct and order on each shelf matters.
\end{enumerate}

\medskip
\solution

\end{document}
