\newcommand{\hmwkTitle}{Homework 7}
\newcommand{\hmwkDueDate}{Tuesday, June 23, 2015}
\newcommand{\hmwkHeaderTitle}{ICCS206 Discrete Mathematic: Homework 7}
\input{hw_template.tex}

\section*{Problem 1: More Counting}
\begin{enumerate}[label=\arabic*.]
    \item Let $x_1,\dots,x_k \in \{1,2,3,\dots\}$. How many solutions satisfy $\sum_{i=1}^k x_i = n$?
    \item Same domain as above. How many solutions satisfy $\sum_{i=1}^k x_i \le n$?
    \item Let $x_1,\dots,x_k \in \{0,1,2,3,\dots\}$. How many solutions satisfy $\sum_{i=1}^k x_i = n$?
    \item Same domain as above. How many solutions satisfy $\sum_{i=1}^k x_i \le n$?
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item we try to frame this as star-and-bars, and as 0 (empty) is allowed, we have to shift the set of $\{1, 2, 3, ...\}$ down by 1, as ${{x^'}_i} = x_i - 1$, so the set would be $\{0, 1, 2, ...\}$ instead; and the summation becomes 
        \[\sum_{i=1}^k {{x^'}_i} = n - k\]
        now, to find all the permutation that satisfy the given criteria, we can apply star-and-bars --- $(n-k)$ stars, k bars --- which gives $\dbinom{(n - k) + k - 1}{k-1} = \boxed{\dbinom{n-1}{k-1}}$ solutions
    \item just like (1.1), we try to frame it as star-and-bars by introduces 0 (emptyness) into the set, but for this one, we have to accounted for the inequality by assigning anouther spot `s` such that $0 \le s \le (n-k)$, and the summation becomes
        \[\sum_{i=1}^k {{x^'}_i} + s= n - k\]
        the star counts remain the same, but as we add another spot/ box, there are now k+1 bars --- which gives $\dbinom{(n - k) + k + 1 - 1}{k + 1 - 1}= \boxed{\dbinom{n}{k}}$ solutions.
    \item this one is just (1.1), but the set already in a proper condition for star-and-bars; so from the summation, just plug it in --- n stars, k bars --- $\boxed{\dbinom{n+k-1}{k-1}}$ solutions.
    \item like (1.2), followed from (1.3), we could rewrite the summation as
        \[\sum_{i=1}^k x_i + s = n \quad \text{ where } 0 \le s \le n\]
        the star counts remain the same, but as we add another spot/ box, there are now k+1 bars --- which gives $\boxed{\dbinom{n+k}{k}}$
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 2}
From 100 classes you record whether AJ Piti ends late and whether it rains:
\[
\begin{array}{c|cc}
& \text{Late} & \text{Not Late} \\
\hline
\text{Rain} & 10 & 15 \\
\text{Not Rain} & 50 & 25
\end{array}
\]
Compute:
\begin{enumerate}[label=\arabic*.]
    \item $P(\text{Rain})$
    \item $P(\text{Late})$
    \item $P(\text{Late} \mid \text{Rain})$
    \item $P(\text{Late} \cap \text{Rain})$
    \item $P(\text{Rain} \mid \text{On Time})$
    \item Are ``Late'' and ``Rain'' independent?
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item $\dfrac{25}{100} = \dfrac{1}{4}$
    \item $\dfrac{60}{100} = \dfrac{3}{5}$
    \item $\dfrac{P(\text{Late})}{P(\text{Rain})} = \dfrac{10}{25} = \dfrac{2}{5}$
    \item $\dfrac{10}{100} = \dfrac{1}{10}$
    \item $\dfrac{P(\text{Rain})}{P(\text{Not Late})} = \dfrac{15}{40} = \dfrac{3}{8}$
    \item no, as \[P(\text{Late $\cap$ Rain}) = \dfrac{1}{10} \ne \left(\dfrac{3}{5}\right)\left(\dfrac{1}{4}\right) = \dfrac{3}{20} = P(\text{Rain}) \cdot P(\text{Late})\]
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 3}
A bag initially contains 2 cola and 3 grape yoyos.
\begin{enumerate}[label=\arabic*.]
    \item Draw two yoyos sequentially without replacement.
    \begin{enumerate}[label=\alph*.]
        \item Probability of one cola and one grape.
        \item Probability of two colas.
        \item Probability the first yoyo is cola.
        \item Given the second yoyo is grape, probability the first is cola.
        \item Given the first is cola, probability the two yoyos are different flavors.
        \item Given the two yoyos are different flavors, probability the first is cola.
        \item Given the two yoyos are the same flavor, probability the first is cola.
    \end{enumerate}
    \item Draw two yoyos simultaneously (without replacement).
    \begin{enumerate}[label=\alph*.]
        \item Probability of one cola and one grape.
        \item Probability of two colas.
    \end{enumerate}
    \item Draw the first yoyo, replace it, then draw a second.
    \begin{enumerate}[label=\alph*.]
        \item Probability of one cola and one grape.
        \item Probability of two colas.
        \item Probability the first yoyo is cola.
        \item Given the second is grape, probability the first is cola.
        \item Given the first is cola, probability the two yoyos are different flavors.
        \item Given the two are different flavors, probability the first is cola.
        \item Given the two are the same flavor, probability the first is cola.
    \end{enumerate}
\end{enumerate}

\medskip
\solution

\begin{enumerate}
    \item let $C$ notates cola, and $G$ notates grape
    \begin{enumerate}
        \item $\dfrac{2}{5}\cdot\dfrac{3}{4} + \dfrac{3}{5}\cdot\dfrac{2}{4} = \dfrac{12}{20} = \boxed{\dfrac{3}{5}}$
        \item $\dfrac{2}{5}\cdot\dfrac{1}{4} = \boxed{\dfrac{1}{10}}$
        \item $\boxed{\dfrac{2}{5}}$
        \item $\dfrac{2}{4} = \boxed{\dfrac{1}{2}}$
        \item there'll be 1 cola, 3 grapes left, so the probability that we get grape is $\boxed{\dfrac{3}{4}}$
        \item $P(CG) = P(GC)$, so $\boxed{\dfrac{1}{2}}$
        \item if the first yoyo gonna be a cola, it must be $CC$, so
            \[\dfrac{P(CC)}{P(CC)+P(GG)} = \dfrac{\frac{1}{10}}{\frac{1}{10}+\frac{3}{10}} = \boxed{\dfrac{1}{4}}\]
    \end{enumerate}

    \medskip
    \begin{center}
    \rule{0.3\linewidth}{0.2pt}
    \end{center}
    \medskip

    \item
    \begin{enumerate}
        \item $\dfrac{\binom{2}{1}\binom{3}{1}}{\binom{5}{2}} = \boxed{\dfrac{3}{5}}$
        \item $\dfrac{\binom{2}{2}}{\binom{5}{2}} = \boxed{\dfrac{1}{10}}$
    \end{enumerate}

    \medskip
    \begin{center}
    \rule{0.3\linewidth}{0.2pt}
    \end{center}
    \medskip

    \item 
    \begin{enumerate}
        \item $2\left(\dfrac{2}{5}\cdot\dfrac{3}{5} \right)= \boxed{\dfrac{12}{25}}$
        \item $\left(\dfrac{2}{5}\right)^2 = \boxed{\dfrac{4}{25}}$
        \item $\boxed{\dfrac{2}{5}}$
        \item the first yoyo got put back anyway, so $\boxed{\dfrac{2}{5}}$
        \item it's just the probability that we got grape for second draw (after put back the first yoyo), so $\boxed{\dfrac{3}{5}}$
        \item $\boxed{\dfrac{1}{2}}$
        \item $\dfrac{P(CC)}{P(CC)+P(GG)} = \dfrac{\frac{4}{25}}{\frac{4}{25}+\frac{9}{25}} = \boxed{\dfrac{4}{13}}$
    \end{enumerate}
\end{enumerate}

\vspace{1em}
\hrulefill

\section*{Problem 4}
Three dice have faces
\[
A = \{2,6,7\},\quad B = \{1,5,9\},\quad C = \{3,4,8\}.
\]
In class we saw $A$ beats $B$, $B$ beats $C$, and $C$ beats $A$ when rolling once. Now roll the chosen die twice and sum the results; compare dice under this rule. Determine which dice beat which.

\medskip
\solution

compute the two-roll outcomes' sum for each die, by brute force, we get:

\[
\begin{aligned}
A&={2,6,7}:\quad 4(1),\ 8(2),\ 9(2),\ 12(1),\ 13(2),\ 14(1) \\
B&={1,5,9}:\quad 2(1),\ 6(2),\ 10(3),\ 14(2),\ 18(1) \\
C&={3,4,8}:\quad 6(1),\ 7(2),\ 8(1),\ 11(2),\ 12(2),\ 16(1)
\end{aligned}
\]

and we just count the scores for the die pairs, from the total of 9 x 9 = 81 (i supposed you could choose to drop the ties, or just count it, as we only see which die (of the pair) got better probability of winning)

% generated by codex cli, i did count it manually as well (as you see in the fractions below)
\begin{tikzpicture}[x=1cm,y=0.9cm]
\node[anchor=west] at (0,1) {A two-roll outcomes (each $1/9$)};
\node (A1) at (0,0) {$A:(2,2)\to 4$};
\node (A2) at (0,-1) {$A:(2,6)\to 8$};
\node (A3) at (0,-2) {$A:(2,7)\to 9$};
\node (A4) at (0,-3) {$A:(6,2)\to 8$};
\node (A5) at (0,-4) {$A:(6,6)\to 12$};
\node (A6) at (0,-5) {$A:(6,7)\to 13$};
\node (A7) at (0,-6) {$A:(7,2)\to 9$};
\node (A8) at (0,-7) {$A:(7,6)\to 13$};
\node (A9) at (0,-8) {$A:(7,7)\to 14$};

\node[anchor=west] at (6,1) {B two-roll outcomes (each $1/9$)};
\node (B1) at (6,0) {$B:(1,1)\to 2$};
\node (B2) at (6,-1) {$B:(1,5)\to 6$};
\node (B3) at (6,-2) {$B:(1,9)\to 10$};
\node (B4) at (6,-3) {$B:(5,1)\to 6$};
\node (B5) at (6,-4) {$B:(5,5)\to 10$};
\node (B6) at (6,-5) {$B:(5,9)\to 14$};
\node (B7) at (6,-6) {$B:(9,1)\to 10$};
\node (B8) at (6,-7) {$B:(9,5)\to 14$};
\node (B9) at (6,-8) {$B:(9,9)\to 18$};

\node[anchor=west] at (10,1);
\node (R1) at (10,0) {$B<A$};
\node (R2) at (10,-1) {$B<A$};
\node (R3) at (10,-2) {$B>A$};
\node (R4) at (10,-3) {$B<A$};
\node (R5) at (10,-4) {$B>A$};
\node (R6) at (10,-5) {$B>A$};
\node (R7) at (10,-6) {$B>A$};
\node (R8) at (10,-7) {$B>A$};
\node (R9) at (10,-8) {$B>A$};

\draw[->] (A2) -- (B1);
\draw[->] (A2) -- (B2);
\draw[->] (A2) -- (B3);
\draw[->] (A2) -- (B4);
\draw[->] (A2) -- (B5);
\draw[->] (A2) -- (B6);
\draw[->] (A2) -- (B7);
\draw[->] (A2) -- (B8);
\draw[->] (A2) -- (B9);

\draw[->] (B1) -- (R1);
\draw[->] (B2) -- (R2);
\draw[->] (B3) -- (R3);
\draw[->] (B4) -- (R4);
\draw[->] (B5) -- (R5);
\draw[->] (B6) -- (R6);
\draw[->] (B7) -- (R7);
\draw[->] (B8) -- (R8);
\draw[->] (B9) -- (R9);
\end{tikzpicture}

by doing boring thing like the illustration above, we will get:
\[
P(B>A)=\frac{1\cdot 0+2\cdot 1+3\cdot 5+2\cdot 8+1\cdot 9}{81}=\frac{42}
{81}.
\]

\[
P(C>B)=\frac{1\cdot 1+2\cdot 3+1\cdot 3+2\cdot 6+2\cdot 6+1\cdot 8}{81}
=\frac{42}{81}.
\]

and

\[
P(A>C)=\frac{1\cdot 0+2\cdot 3+2\cdot 4+1\cdot 6+2\cdot 8+1\cdot 8}{81}
=\frac{44}{81}.
\]

so the cycle reverses as $\boxed{B\succ A,\ C\succ B,\ A\succ C}$

\vspace{1em}
\hrulefill

\section*{Problem 5}
Three identical boxes each contain two yoyos. Two boxes have one cola and one grape; the third has two colas. Play the game:
\begin{enumerate}[label=\arabic*.]
    \item First we pick a box.
    \item Then we draw a yoyo for that box.
    \item If it is a grape yoyo, we select the box again.
    \item If it is a cola yoyo then we can continue the game.
    \item We then draw another yoyo from the box. If the next yoyo is cola AJ wins. If the next yoyo is grape, you win.
\end{enumerate}
Is this a fair game? Justify.

\medskip
\solution

it's a fair game, as in stage (4), we already pick up a cola yoyo, so the best strategy that doesn't rely on guessing/ luck is to pick the remaining yoyo from the same box we just draw a cola yoyo from --- which there are 2 outcomes 

\begin{itemize}
    \item another yoyo is also a cola, and we lost --- this got a probability of $\dfrac{1}{3}$, as there's only one box without a grape yoyo
    \item another yoyo is a grape yoyo, this got a probability of $\dfrac{2}{3}$, but we have to accounted for our chance of picking up a grape yoyo for our first draw, which is $\dfrac{1}{2}$ --- so the probability is $\dfrac{2}{3} \cdot \dfrac{1}{2} = \dfrac{1}{3}$
\end{itemize}

as the probility of getting a cola yoyo on the second draw, and getting a grape yoyo on the second draw are the same, we can conclude this is a fair game.

\vspace{1em}
\hrulefill

\section*{Problem 6}
AJ Piti will give a rare banana yoyo to Bossy, Tow$+$, or John, each equally likely. AJ secretly selects the recipient. If Bossy asks for one name who will not get the yoyo, AJ responds as follows:
\begin{itemize}
    \item If Bossy does not get the yoyo, AJ names the other person who also does not get it.
    \item If Bossy does get it, AJ flips a fair coin to name John or Tow$+$.
\end{itemize}
AJ tells Bossy that John will not get the yoyo. Bossy concludes his chance is $1/2$, while Tow$+$ claims it remains $1/3$ for Bossy and $2/3$ for himself. Determine the actual probabilities for Bossy and Tow$+$ given AJ's statement. (Hint: draw a probability tree.)

\medskip
\solution

it can't be 50:50, as it felt like the way AJ responds favours the guy that doesn't gets mentioned, at first, say (Bossy, Tow+, John) $\equiv$ (B, T, J), we know that

\[P(B)=P(T)=P(J)=\tfrac{1}{3}\]

let $R_j$ be AJ responds as John, we'll get 

\[
P(R_j\mid B)=\tfrac{1}{2},\quad P(R_j\mid T)=1,\quad P(R_j\mid J)=0,
\]

as, if Tow+ is the winner, AJ would have to say John --- that's probability of 1.0 of AJ saying John; and if Bossy is the winner, there's a probability of $\dfrac{1}{2}$ that AJ will say Tow+(another half is for John) --- that's only $\dfrac{1}{2}$ of assurance that Bossy will win if AJ says John. 

\vspace{1em}

so, the change of Bossy winning after AJ responds as John is

\[
P(B\mid R_j)=\frac{\frac{1}{2}\cdot\frac{1}{3}}{\frac{1}{2}}=\boxed{\tfrac{1}{3}},\qquad
\]

and Tow+'s change of winning after AJ responds with John is

\[
P(T\mid R_j)=\frac{1\cdot\frac{1}{3}}{\frac{1}{2}}=\boxed{\tfrac{2}{3}}.
\]

so Tow$+$ is correct.

\end{document}
